\thetitle

### 2020-10-20 sun_phases/spin_qudit_tomo.tex
added a tiny bit to the intro added an appendix deriving the translation between solutions to the signal recovery problem in different bases

\begin{abstract}
  We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
  By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
  We then provide an explicit optimized tomographic protocol, as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.
  [\red{Note: there are also a few other things that I want to look into, such as (1) adaptive tomography protocols, (2) proofs about something like the number of measurements necessary to reconstruct an arbitrary state to a given precision, (3) a simplified tomography procedure for measuring pure states, (4) the trade-off between measurement precision and the \# of measurement axes (for a fixed total number of measurements).]}
\end{abstract}

@@

In this work, we provide a new and simple framework for performing spin qudit tomography.
We begin by mapping this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere, thereby importing a host of existing literature and machinery that has been developed for this task \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.

[\red{finish intro, provide outline of the paper}]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin transition operators and rotation matrices}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin tomography as signal recovery on the sphere}

Our goal is to reconstruct an arbitrary state of a spin qudit from measurements of spin projection onto different quantization axes.
We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis determined by a point $\v v$ on the 2-sphere $\Omega_2$.
For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v \ell,0}}_\rho\equiv\tr\p{\rho T_{\v v\ell,0}}$.

@@

Moreover, any solution to the full recovery problem with a particular set of basis functions, such the spherical harmonics $Y_{\ell m}$, can be readily translated into a solution with a different set of basis functions, such as the adjoint rotation matrix elements $\D^\ell_{m,0}$.
This translation enables importing a host of existing signal recovery algorithms \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
Specifically, given a set of points $V\subset\Omega_2$ and coefficients $C_{\ell m}\p{\v v}$ that satisfy Eq.~\eqref{eq:full_recovery} with basis functions $Y_{\ell m}$, we can construct new coefficients $\tilde C_{\ell m}\p{\v v}$ that satisfy the same condition with basis functions $\D^\ell_{m,0}$ via
\begin{align}
  \tilde C_{\ell m} \equiv \sum_n G^\ell_{mn} C_{\ell n},
\end{align}
where
\begin{align}
  G^\ell_{mn} \equiv \f{2\ell+1}{4\pi}
  \int_{\Omega_2} \d\Omega_2\p{\v v}
  \D^\ell_{m,0}\p{\v v}^* Y_{\ell n}\p{\v v},
\end{align}
is an inner product of the functions $\D^\ell_{m,0}$ and $Y_{\ell n}$ [\red{does this need to be derived in its own Appendix?}].

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tomographic protocol and numerical experiments}

@@

>>>>

\begin{abstract}
  We consider the task of performing quantum state tomography on a $d$-state spin qudit, using only measurements of spin projection onto different quantization axes.
  By mapping this quantum tomography problem onto the classical problem of signal recovery on the sphere, we prove that full reconstruction of arbitrary mixed states is possible, and requires a minimum number of measurement axes, $r_d$, that is bounded by $2d-1\le r_d\le d^2$.
  We then provide an explicit optimized tomographic protocol, as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.
  [\red{Note: there are also a few other things that I want to look into, including (1) a simplified tomography procedure for measuring pure states, (2) trade-offs between measurement precision and the \# of measurement axes (for a fixed total number of measurements; these will likely be numerical experiments), (3) proofs about something like the number of measurements necessary to reconstruct an arbitrary state to a given precision, and (4) adaptive tomography protocols.]}
\end{abstract}

@@

In this work, we provide an efficient framework for performing spin qudit tomography using only measurements of spin projection onto different spatial axes.
We first introduce, in Section \ref{sec:transition_ops}, a basis of operators that have a simple physical interpretation, and transform nicely under rotations in 3D space, thereby respecting the symmetries of the tomography task at hand.
We then map this quantum tomography problem onto the well-studied classical problem of signal recovery on the sphere in Section \ref{sec:signal_recovery}, thereby importing a host of existing literature and machinery that has been developed for this task \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
In particular, this mapping allows us to bound the minimum number of measurement axes necessary to perform full tomography on a $d$-level spin qudit, $r_d$, by $2d-1\le r_d\le d^2$.
We provide an example of an optimized tomography protocol in Section \ref{sec:protocol}, as well as numerical experiments to show that $r_d=2d-1$ for $d\le20$.

[\red{finish intro}]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin transition operators and rotation matrices}
\label{sec:transition_ops}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Spin tomography as signal recovery on the sphere}
\label{sec:signal_recovery}

Our goal is to reconstruct an arbitrary state of a spin qudit from measurements of spin projection onto different quantization axes.
We are thus nominally restricted to measuring projectors $\Pi_{\v v\mu} \equiv \op{\mu_{\v v}}$, where $\ket{\mu_{\v v}}\equiv R\p{\v v}\ket{\mu}$ is a state with spin projection $\mu$ onto a measurement axis determined by a point $\v v$ on the 2-sphere $\Omega_2$.
For any fixed axis $\v v$, the sets $\set{\Pi_{\v v\mu}}$ and $\set{T_{\v v\ell,0}}$ are both orthonormal bases for the space of operators that are diagonal in the basis $\set{\ket{\mu_{\v v}}}$.
By a simple change of operator basis, measuring the projectors $\set{\Pi_{\v v\mu}}$ is therefore equivalent to measuring the transition operators $\set{T_{\v v\ell,0}}$, and provides data on the expectation values $\bk{T_{\v \ell,0}}_\rho$, where $\bk{X}_\rho\equiv\tr\p{\rho X}$.

@@

Moreover, any solution to the full recovery problem with a particular set of basis functions, such the spherical harmonics $Y_{\ell m}$, can be readily translated into a solution with a different set of basis functions, such as the adjoint rotation matrix elements $\D^\ell_{m,0}$.
This translation enables importing a host of existing signal recovery algorithms \cite{mcewen2011novel, mcewen2011sampling, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
Specifically, given a set of points $V\subset\Omega_2$ and coefficients $C_{\ell m}\p{\v v}$ that satisfy Eq.~\eqref{eq:full_recovery} with basis functions $Y_{\ell m}$, we can construct new coefficients $\tilde C_{\ell m}\p{\v v}$ that satisfy the same condition with basis functions $\D^\ell_{m,0}$ via
\begin{align}
  \tilde C_{\ell m} \equiv \sum_n G^\ell_{mn} C_{\ell n},
\end{align}
where
\begin{align}
  G^\ell_{mn} \equiv \f{2\ell+1}{4\pi}
  \int_{\Omega_2} \d\Omega_2\p{\v v}
  \D^\ell_{m,0}\p{\v v}^* Y_{\ell n}\p{\v v},
\end{align}
is an inner product of the functions $\D^\ell_{m,0}$ and $Y_{\ell n}$ (see Appendix \ref{sec:basis_change}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Tomography protocol and numerical experiments}
\label{sec:protocol}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Changing bases for the signal recovery problem}
\label{sec:basis_change}

Here we show how one solution to the problem of signal recovery on the sphere with a particular basis of functions (such as the spherical harmonics $Y_{\ell m}$) can be translated into a solution with a different basis (such as the adjoint rotation matrix elements $\D^\ell_{m,0}$).
Consider a square-integrable function on the sphere, $f\in L^2\p{\Omega_2}$, with the harmonic expansion
\begin{align}
  f\p{\v v} = \sum_{\ell,m} f_{\ell m} B_{\ell m},
\end{align}
where $f_{\ell m}\in\CC$ is a coefficient and $\set{B_{\ell m}}$ is a complete basis for $L^2\p{\Omega_2}$.
Without loss of generality, we will assume that for a fixed degree $\ell$ the functions $\set{B_{\ell m}}$ span the degree-$\ell$ subspace of $L^2\p{\Omega_2}$, and that these functions satisfy orthogonality relations analogous to that in Eq.~\eqref{eq:ortho_sphere}.
A solution to the signal recovery problem amounts to a set of points $V=\set{\v v}\subset\Omega_2$ and coefficients $C_{\ell m}\p{\v v}$ for which
\begin{align}
  f_{\ell m} = \sum_{\v v} C_{\ell m}\p{\v v} f\p{\v v}.
  \label{eq:recovery_solution}
\end{align}
Consider now a different basis $\set{\tilde B_{\ell m}}$ for $L^2\p{\Omega_2}$, with which
\begin{align}
  f\p{\v v} = \sum_{\ell, m} \tilde f_{\ell m} \tilde B_{\ell m},
\end{align}
where we make the same general assumptions about $\set{\tilde B_{\ell m}}$ that we made for $\set{B_{\ell m}}$.
Substituting the harmonic expansion of $f\p{\v v}$ in the original basis $\set{B_{\ell m}}$, we find that
\begin{align}
  \tilde f_{\ell m} = \int_{\Omega_2} \d\Omega_2\p{\v v}
  \tilde B_{\ell m}\p{\v v}^* f\p{\v v}
  = \sum_n G^\ell_{mn} f_{\ell n},
  &&
  G^\ell_{mn} \equiv \int_{\Omega_2} \d\Omega_2\p{\v v}
  \tilde B_{\ell m}\p{\v v}^* B_{\ell n}\p{\v v},
  \label{eq:basis_change}
\end{align}
where integration is performed over all $\v v\in\Omega_2\simeq[0,2\pi]\times[0,\pi]$ with integration measure $\d\Omega_2\p{\v v} \equiv \d\alpha_{\v v} \d\beta_{\v v} \sin\beta_{\v v}$.
Note that the inner product between functions $\tilde B_{\ell m}\p{\v v}$ and $B_{kn}\p{\v v}$ for $k\ne\ell$ vanishes because these functions live in orthogonal subspaces of $L^2\p{\Omega_2}$ \cite{freeden2008spherical}.
Substituting Eq.~\eqref{eq:recovery_solution} into Eq.~\eqref{eq:basis_change}, we find that
\begin{align}
  \tilde f_{\ell m}
  = \sum_n G^\ell_{mn} \sum_{\v v} C_{\ell n}\p{\v v} f\p{\v v}
  = \sum_{\v v} \tilde C_{\ell m}\p{\v v} f\p{\v v},
  &&
  \tilde C_{\ell m}\p{\v v} \equiv \sum_n G^\ell_{mn} C_{\ell n}\p{\v v}.
\end{align}
The coefficients $\tilde C_{\ell m}\p{\v v}$ thus solve the signal recovery problem with the new basis $\set{\tilde B_{\ell m}}$ for $L^2\p{\Omega_2}$.
